\documentclass{amsart}
\usepackage{amsmath,amssymb,graphicx,subfig,mathrsfs}
\usepackage{hyperref}
\newcommand{\norm}[1]{\left\Vert #1 \right\Vert}
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
\begin{document}
\title{$\log\sin x$}
\author{Jordan Bell}
\email{jordan.bell@gmail.com}
\address{Department of Mathematics, University of Toronto, Toronto, Ontario, Canada}
\date{\today}

\maketitle

That 
\[
\int_0^{2\pi} \log|1-e^{i\theta}| d\theta = 0.
\]
is used in the proof of Jensen's formula.

Euler \cite[\S\S 35--37]{E393} E393,  \cite[\S 21]{E432}, \cite{E499} E499
 
 Class number formula
 
Koyama and Kurokawa \cite{koyama}

Mathematical Fallacies, Flaws and Flimflam
 By E. J. Barbeau
 p. 105
 
 complex analysis: Rudin, Real and Complex, p.~307
 
 Theorie analytique, Fourier, Chapter III, Sect. III, p. 149, no. 183, Fourier series for logsin
 
 sect. 192 of Introductio in analysin, chap. 11
  
 cotangent series: E128
 
 Littrow 
 
 Oeuvres completes de N. H. Abel, tome II, p. 92
 Sur quelques integrales definies
 
 Archiv der Mathematik und Physik, Volume 4
Johann August Grunert
113-119
1844

Isaac Todhunter
pp. 65--66
\S 51
A Treatise on the Integral Calculus and Its Applications with Numerous Examples
1874
Macmillan and Co.
Fourth ed.

Methoden der mathematischen Physik, Volume 1
p. 88

George Chrystal
367
Part 2

Reihentheorie
 By Jakob Bernoulli
 Band 4
 
 \cite[p.~8]{MR1700272}
 
 Examples of the processes of the differential and integral calculus
 By Duncan Farquharson Gregory
 484-485, cites Ellis
 
 The Cauchy Method of Residues, Volume 2: Theory and Applications
 By D. S. Mitrinovic, Jovan D. Ke?ki?
 p. 68
 
 An Introduction to the Theory of Infinite Series
 By Thomas John I'Anson Bromwich
 p. 467
 
 N. Nielsen. Handbuch der theorie der Gammafunction. Druck und Verlag von B.G. Teubner, Leipzig, 1906
 
 Mémoires de l'Académie impériale des sciences, littérature et Turin, 1811-1812, vol 20, Georges Bidone, 231-344
 
 De calculo integralium exercitatio mathematica, Pietro Ferroni
 
 Vorlesungen über Geschichte der Trigonometrie, Volume 2, Anton Braunmuhl
 
 Foundations of Hyperbolic Manifolds, second ed., John Ratcliffe, references, p. 506 (Clausen function and Lobachevsky function)
 
 Traité élémentaire de calcul différentiel et de calcul intégral ..., Volume 2
 By Silvestre F. Lacroix
 p. 350
 
 A Synopsis of Elementary Results in Pure Mathematics: Containing ...
 By George Shoobridge Carr
 
 Die Euler'schen integrale bei unbeschränkter variabilität der arguments
 By Hermann Hankel
 
 \nocite{*}

\bibliographystyle{amsplain}
\bibliography{logsin}

\end{document}